---
title: "R Notebook"
output: html_notebook
---


```{r}
# Read in the data
hrdata <- read.csv("HR-Employee-Attrition.csv")
```
```{r}
dim(hrdata)
```
```{r}

df_left <- hrdata[c("Attrition", "MonthlyIncome", "Gender", "JobRole", "PercentSalaryHike", "YearsAtCompany", "YearsInCurrentRole","YearsSinceLastPromotion", "YearsWithCurrManager", "DistanceFromHome")]
df_left_yes <- df_left[df_left$Attrition == "Yes",] 
df_left_yes
```
```{r}

install.packages('writexl')
library(writexl)

write_xlsx(df_left_yes, 'C:\\Users\\jfauxinu\\Documents\\hr_analysis_module8\\data.xlsx')

```
			

```{r}
# correlation matrix for selected columns to see if the demographics correlate
cor(
hrdata[ , c("Age", "DailyRate", "DistanceFromHome", "Education", "HourlyRate", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", "TotalWorkingYears", "TrainingTimesLastYear")]
)
```
#### Strong positive correlation
##### Monthly Income and Total Working Years has a .77 correlation - more years worked more monthly income earned
 
##### Age and Total Working Years has a .68 correlation  - more years worked employees are older

##### Age and Monthly Income has a .5 correlation - generally the older people are the more their monthly income

#### Positive correlation

##### Age and NumberCompaniesWorked has a .30 correlation

#####  Age and Education has a .21

```{r}
pairs(~MonthlyIncome+Age+TotalWorkingYears+Education,data = hrdata,
main = "Scatterplot Matrix")
```
#### Strong positive correlation

##### Monthly Income and Age 

##### Monthly Income and Total Working Years

##### Age and Total Working Years

```{r}
# Install  packages - only have to do this once
#install.packages("ggplot2")
#install.packages("GGally")

# Load ggplot2
library("ggplot2")
```

```{r}
# Load GGally
library(GGally)

```


```{r}
# add Gender to see the correlations
ggpairs(hrdata, columns = c('MonthlyIncome', 'Age', 'TotalWorkingYears', 'Education', 'Gender'),
        aes(color = Gender),
        columnLabels = c("Monthly Income", "Age", "Total Working Years", "Education", "Gender"))

```
#### Strong positive correlation
##### Monthly Income and Total Working Years has a .773 correlation - more years worked more monthly income earned
####                                    Female has .761 correlation females earn less
####                                    Male has   .781 males correlation earn more                       
 
#####  Total Working Years and Age has a .680 correlation  - more years worked employees tend to be older
####                        Female has a .704 correlation  - females have worked more years
####                        Male  has  a .663 correlation  - males have worked less years

##### Age and Monthly Income has a .498 correlation - generally the older people are the more their monthly income
####                  Female has a .505 correlation - females are slightly older
####                  Male has a   .482 correlation - males are slightly younger


```{r}
# boxplot of the age distribution of employees who were laid off (yes) and those who were not (no)
boxplot(Age~Attrition,data=hrdata, main= "Who Got Laid Off", xlab="Attrition", ylab="Age",
        col = "orange", border = "brown", horizontal = TRUE, notch = TRUE)

```

```{r}
yes_age <- hrdata[(hrdata$Attrition == "Yes"),'Age']
no_age <- hrdata[(hrdata$Attrition != "Yes"),'Age']
```


```{r}
t.test(yes_age, no_age)
```
Assigned to investigate claim that the last layoffs impacted older employees - age discrimination
The p value is less than 0.05, there is a statistically significant difference between the two samples, but employees that were affected were actually younger than those who remained on the job. Reject the null hypothesis (not based on age)

```{r}

# boxplot of the employee number distribution of employees who were laid off 
boxplot(EmployeeNumber~Attrition,data=hrdata, main= "Who Got Laid Off", xlab="Attrition", ylab="EmployeeNumber",
        col = "indianred1", border = "slateblue", horizontal = TRUE, notch = TRUE)
```

From the boxplot it seems some newer hires were not impacted, but on tail end it seems the numbers were smaller.


```{r}
yes_emp_num <- hrdata[(hrdata$Attrition == "Yes"),'EmployeeNumber']
no_emp_num <- hrdata[(hrdata$Attrition != "Yes"),'EmployeeNumber']
```


```{r}
t.test(yes_emp_num, no_emp_num)
```
Another claim has been filed - this time need to investigate if recently hired employees were impacted more than senior employees (H0).
This time the p value is much larger than 0.05. Employee Number does not have a statistically significant relationship with the Attrition.
Can't say with confidence that newer employees were not impacted more than senior employees. Accept or fail to reject the Null Hypothesis.

```{r}
model1 = lm(MonthlyIncome ~ Age, data = hrdata)
summary(model1)
```
#### Linear regression model that predicts the MonthlyIncome based upon Age
#### p is smaller than 0.05, which is statistically signifcant.
#### The R-squared value is .2479, the model explains 
#### close to 25% of variance in the MonthlyIncome is due to age. 

```{r}
model2 = lm(MonthlyIncome ~ Age + TotalWorkingYears, data = hrdata)
summary(model2)
```
### The p-value is still smaller than 0.05. The R-squared value is increased to 0.5988, 
### so model2 explains that close to 60% of the variance in the MonthlyIncome 
### and TotalWorkingYears by due to using the age variable.

```{r}
hist(residuals(model1), col = "gray")
```
### Checking Assumptions of the Model
#### The distribution of model residuals should be approximately normal.
###  Above the distribution is slightly right skewed.
###  Belwo the distribution is normal
```{r}
hist(residuals(model2), col = "blue")
```

```{r}
#create fitted value vs residual plot
plot(fitted(model2), residuals(model2))

#add horizontal line at 0
abline(h = 0, lty = 2)
```
The variance of the residuals should be consistent for all observations.



